Importing libraries

```{r, echo=TRUE}
my_packages <- c("opendatatoronto", "dplyr", "ggplot2", "tidyr")
# For each of the 3 packages, check whether each one exists, and install if not
for (pkg in my_packages) {
  
  if (!pkg %in% installed.packages()) { 
    # If package is not installed, install it:
    install.packages(pkg)
  } 
}
# Importing libraries
library(opendatatoronto)
library(dplyr)
library(tidyr)
library(ggplot2)
```

## Finding the data

To find a dataset you like, you have a couple options:

1. Search for a dataset manually via the [Toronto Open Data Portal](https://open.toronto.ca/)
2. Use the `search_packages()` function

Here we'll use the `search_packages()` function to keep this document self-contained.

```{r, echo= TRUE}
# Search packages (this returns a table)
wellbeing_packages <- search_packages("Wellbeing")
# Print the table
wellbeing_packages
```
Get demographic package as the main package
```{r, echo= TRUE}
wellbeing_demographics_package<- wellbeing_packages %>%
  filter(title == "Wellbeing Toronto - Demographics: NHS Indicators") # Only keep the row(s) where the title is "Wellbeing Toronto - Demographics: NHS Indicators"
```

Get demographic dataset from the package
```{r, echo= TRUE}
demographic_indicators2011 <- wellbeing_demographics_package %>% # Start with the package 
  list_package_resources() %>% 
  filter(name == "demographics-nhs-indicators-2011") %>% 
  get_resource()
```

Get Housing package for neighbourhood name retrieval 
```{r, echo= TRUE}
wellbeing_housing_package<- wellbeing_packages %>%
  filter(title == "Wellbeing Toronto - Housing") 
```

Get Housing dataset from the package, choosing a specific sheet
```{r, echo= TRUE}
housing_data <- wellbeing_housing_package %>% # Start with the package 
  list_package_resources() %>% # List the resources in the package
  filter(name == "wellbeing-toronto-housing") %>% 
  get_resource()
housing_data <- housing_data$RawDataRef_2011
```

Slice the demographic dataset
```{r, echo=TRUE}
demographic_indicators2011 <- demographic_indicators2011 %>%
  dplyr::slice(2:142)
demographic_with_names <-demographic_indicators2011
```

Rename wanted columns
```{r, echo=TRUE}
demographic_with_names <- demographic_with_names %>% 
  rename(
    'hood' = 'Hood#'
    )
demographic_with_names <- demographic_with_names %>% 
  rename(
    'median_household_income' = 'Median household total income $'
    )
demographic_with_names <- demographic_with_names %>% 
  rename(
    'total_pop' = 'Total Population'
    )
demographic_with_names <- demographic_with_names %>% 
  rename(
    'no_certificate' = 'No certificate, diploma or degree'
    )
demographic_with_names <- demographic_with_names %>% 
  rename(
    'college' = 'College, CEGEP or other non-university certificate or diploma'
    )
demographic_with_names <- demographic_with_names %>% 
  rename(
    'bachelor' = 'University certificate or diploma below bachelor level'
    )
demographic_with_names <- demographic_with_names %>% 
  rename(
    'total_visible_minority' = 'Total visible minority population'
    )
demographic_with_names <- demographic_with_names %>% 
  rename(
    'not_visible_minority' = 'Not a visible minority'
    )

```

Add neighbourhood column (neighbourhood names)
```{r, echo=TRUE}
# neighbourhood
demographic_with_names <- 
  mutate(demographic_with_names, neighbourhood =
           housing_data$Neighbourhood)
```

Add percentage columns for education levels and visible minority profile (maybe a for loop)
```{r, echo=TRUE}
# No_Certificate_pt
demographic_with_names <- 
  mutate(demographic_with_names, no_certificate_pt =
           as.numeric(demographic_with_names$no_certificate)/as.numeric(demographic_with_names$total_pop))

# College
demographic_with_names <- 
  mutate(demographic_with_names, college_pt =
           as.numeric(demographic_with_names$college)/as.numeric(demographic_with_names$total_pop))

# Bachelor
demographic_with_names <- 
  mutate(demographic_with_names, bachelor_pt =
           as.numeric(demographic_with_names$bachelor)/as.numeric(demographic_with_names$total_pop))

# Total_Visible_Mino
demographic_with_names <- 
  mutate(demographic_with_names, total_visible_minority_pt =
           as.numeric(demographic_with_names$total_visible_minority)/as.numeric(demographic_with_names$total_pop))

# Not_Visible_Mino
demographic_with_names <- 
  mutate(demographic_with_names, not_visible_minority_pt =
           as.numeric(demographic_with_names$not_visible_minority)/as.numeric(demographic_with_names$total_pop))
```

Select wanted columns and filter them
```{r, echo=TRUE}
demographic_with_names <- transform(demographic_with_names, median_household_income = as.numeric(median_household_income))

income <-demographic_with_names$median_household_income
target <- c(max(income), min(income))

demographic_with_names <- demographic_with_names %>%
  select(
    hood,
    neighbourhood,
    total_pop,
    median_household_income,
    no_certificate_pt,
    college_pt,
    bachelor_pt,
    total_visible_minority_pt,
    not_visible_minority_pt,
  ) %>%
  filter(median_household_income %in% target)
demographic_with_names
```
To do: Add a row for Toronto average
