---
title: "Demography of Education in Toronto's neighbourhoods with Highest and Lowest Income"
date: "23/01/2020"
output:
  pdf_document: default
authors: Ke-Li Chiu, Diego Mamanche Castellanos
abstract: what was done, what was found, and why this matters. Descriptive, no prediction
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r, echo=FALSE, include=FALSE}
# installing packages
my_packages <- c("opendatatoronto", "dplyr", "ggplot2", "tidyr")
# For each of the 3 packages, check whether each one exists, and install if not
for (pkg in my_packages) {
  
  if (!pkg %in% installed.packages()) { 
    # If package is not installed, install it:
    install.packages(pkg)
  } 
}
# Importing libraries
library(opendatatoronto)
library(dplyr)
library(tidyr)
library(ggplot2)
```

# Wellbeing Toronto - Demographics: NHS Indicators
The data set package we delved into is *Wellbeing Toronto - Demographics: NHS Indicators*, which provides information from the 2011 National Household Survey (NHS) and presents a profile of Toronto. Data is broken down by the 140 neighbourhoods that make up the City of Toronto. With this data set package, we first identify the two neighbourhoods that have the highest income and the lowest income and then analysize the education demographics in these two neighbourhoods to understand difference of education attainment between these two neighbourhood. We also identify the median household income and the average education levels of the city to be compared with the two neighbourhoods.

There are two data sets in this package in which we combine into one single data set: _Demographics NHS Indicators 2011_ and _Education NHS Indicators 2011_. _Demographics NHS Indicators 2011_ shows the following indicators: population breakdown by ethnicity, household income, education level, employment, age and sex. We extract the indicator of household income to identify the two neighbourds that has the highest household income and the one with the lowest. Although this data set provides data for education levels, the indicators are not completed since the data for education level higher than bachlor is not provided. To obtain the indicators of all education levels, we use _Education NHS Indicators 2011_, which provides complete education level indicators ranged from no-certification to postgraduate-level certification.  

The neighbourhoods in the package are coded by numbers. To to retrieve the names of neighbourhood, we also include the housing data set from *Wellbeing Toronto - Housing* package. We incorporate the housing data set striclty for extracting the names of the neighbourhoods; no housing indicators are used in our analysis. 

```{r, echo= FALSE}
# Search packages (this returns a table)
wellbeing_packages <- search_packages("Wellbeing")
# Print the table
#wellbeing_packages
```


```{r, echo= FALSE}
### Filtering the wellbeing demographic dataset
wellbeing_demographics_package<- wellbeing_packages %>%
  filter(title == "Wellbeing Toronto - Demographics: NHS Indicators") # Only keep the row(s) where the title is "Wellbeing Toronto - Demographics: NHS Indicators"
```

## Demographics NHS Indicators 2011 data set
_Demographics NHS Indicators 2011_ shows the profile of each neighbourhood regarding population breakdown by ethnicity, household income, education level, employment, age and sex. The indicators we are interested in are *Median household total income* and *Median after-tax household income*. These two columns show the median household income of every neighbourhood before and after tax in Canadian dollars. 
```{r, echo= FALSE}
# Getting the information dataset called "demographics-nhs-indicators-2011"
demographic_indicators2011 <- wellbeing_demographics_package %>% # Start with the package 
  list_package_resources() %>% # List the resources in the package
  filter(name == "demographics-nhs-indicators-2011") %>% # Only keep the resource we want 
  get_resource()
head(demographic_indicators2011[,c(1,47:48)]) # Show a preview of the data set
```



```{r, echo=FALSE}
#Cleaning the column names using janitor::clean_names
demographic_indicators2011 <-
  janitor::clean_names(demographic_indicators2011)
```



```{r, echo= FALSE}
#Selecting columns needed
income_demographic2011 <- subset(demographic_indicators2011, select = c("hood_number","total_population", "median_after_tax_household_income","median_household_total_income", 'total_visible_minority_population', 'not_a_visible_minority'))
names(income_demographic2011)[names(income_demographic2011) == "hood_number"] <- "ID"
#income_demographic2011  <- as.data.frame(income_demographic2011)
income_demographic2011 <- income_demographic2011[income_demographic2011$ID != 'ID',]       
#income_demographic2011
```

## Education NHS Indicators 2011

_Education NHS Indicators_ is provided for the population of age 15 and over living in private households. The data set presents the distrubution of people's most advanced education level in each neighbourhood. The levels we selected in the analysis are presented as follow.    

```{r, echo= FALSE}
education_indicators2011 <- wellbeing_demographics_package %>% # Start with the package 
  list_package_resources() %>% # List the resources in the package
  filter(name == "education-nhs-indicators-2011") %>% # Only keep the resource we want 
  get_resource()
names(education_indicators2011[,c(3,4,7,10,11)])
```




```{r, echo= FALSE}
#Cleaning the column names using janitor::clean_names
education_indicators2011 <-
  janitor::clean_names(education_indicators2011)
#head(education_indicators2011[,1:4])
```


```{r echo= FALSE}
#Extracting the neighbourhood code from the column "geography"
edu_id <- extract(education_indicators2011, geography, into = "ID" , regex = "([(][0-9]+[)])")
edu_id <- extract(edu_id, ID, into = "ID" , regex = "([0-9]+)")
#head(edu_id[,1:4])
```


## Wellbeing Toronto - Housing

*Wellbeing Toronto - Housing* data set is used in order to extract the names of the neighbourhood. The indicators of housing are not in the scope of our analysis.

# ===  Writing below this section is not yet started ===
```{r, echo= FALSE}
###Filterint the housing dataset in order to extract the names of each neighbourhood.
wellbeing_housing_package <- wellbeing_packages %>%
  filter(title == "Wellbeing Toronto - Housing") 
#Getting the information dataset called "wellbeing-toronto-housing"
housing_indicators2011 <- wellbeing_housing_package %>% # Start with the package 
  list_package_resources() %>% # List the resources in the package
  filter(name == "wellbeing-toronto-housing") %>% # Only keep the resource we want 
  get_resource()
housing_indicators2011 <- housing_indicators2011$RawDataRef_2011
head(housing_indicators2011$Neighbourhood)
```

```{r echo= FALSE}
#Cleaning the column names using janitor::clean_names, and selecting columns needed
housing_indicators2011 <- 
  janitor::clean_names(housing_indicators2011)
housing_indicators2011 <- subset(housing_indicators2011, select = c("neighbourhood_id", "neighbourhood"))
names(housing_indicators2011)[names(housing_indicators2011) == "neighbourhood_id"] <- "ID"
#housing_indicators2011
```


## Merging all datasets

Explanation

```{r echo= FALSE}
merged_df <- merge(housing_indicators2011, income_demographic2011, by = 'ID')
merged_df <- merge(merged_df, edu_id, by = 'ID')
merged_df <- merged_df[,1:16]
#head(merged_df)
```

```{r echo=FALSE}
#CLEANING THE MERGED DATASET
#Calculate the median for column "median_after_tax_household_income"
#Note: In this case, the median is better due to outliers
  
toronto_median_income <- median(as.numeric(income_demographic2011$median_after_tax_household_income))
toronto_median_income
#Calculate mean for the rest of the columns
toronto_other_columns <-  mutate_all(merged_df[,4:16], as.numeric)
toronto_other_columns <- colMeans(toronto_other_columns)
toronto_other_columns <- as.data.frame(toronto_other_columns)
toronto_other_columns <- round(toronto_other_columns[,1])
toronto_other_columns
neighbourhoods <- c(merged_df[which.max(merged_df$median_after_tax_household_income),]$neighbourhood, merged_df[which.min(merged_df$median_after_tax_household_income),]$neighbourhood)
#Create graph2 dataframe
filtered_dataset <- filter(merged_df, neighbourhood %in% c(neighbourhoods))
filtered_dataset <- add_row(filtered_dataset)
#Assign values to third row
filtered_dataset[3,2:3] <- c('Toronto', toronto_median_income)
filtered_dataset[3,4:16] <- toronto_other_columns
filtered_dataset
filtered_dataset <- mutate(filtered_dataset, less_than_diploma = high_school_diploma_or_equivalent + 
apprenticeship_or_trades_certificate_or_diploma + 
college_cegep_or_other_non_university_certificate_or_diploma)
#Rename columns
filtered_dataset <- 
  rename(filtered_dataset, 
    median_income = median_household_total_income,
    median_income_after_tax = median_after_tax_household_income,
    no_high_school = no_certificate_diploma_or_degree,
    above_bachelor = university_certificate_diploma_or_degree_at_bachelor_level_or_above,
    university_diploma = university_certificate_or_diploma_below_bachelor_level
    )
filtered_dataset <- mutate(filtered_dataset, total_population = (    no_high_school + above_bachelor + university_diploma + less_than_diploma + bachelors_degree ))
#head(filtered_dataset)
```


#Create Graph1 dataframe 
Median household income
Finally, the median total income is presented in a bar graph comparing the median for that neighbourhood as compared to the median for the entire city of Toronto.

```{r echo= FALSE}
#Create the dataframe graph1_df to plot the values
graph1_df <- select(filtered_dataset, neighbourhood, median_income_after_tax, median_income)
#Reshape the dataset
graph1_df <- gather(graph1_df, "income_type", "income_value", -1)
#graph1_df
```


Plotting the graph

```{r echo= FALSE}
ggplot(graph1_df, aes(x = neighbourhood, y = as.numeric(income_value), fill = income_type)) + 
  # Specify that we want a bar graph:
  geom_bar(stat="identity", position=position_dodge()) + 
  # Add titles and axis labels:
  labs(title = "neighbourhoods with the Highest and Lowest median income, \nand Toronto median income", 
       x = "neighbourhood", 
       y = "Income") 
```

## Create graph2 dataframe

```{r echo= FALSE}
#Create the dataframe graph1_df to plot the values
graph2_df <- select(filtered_dataset, neighbourhood, no_high_school, less_than_diploma, university_diploma, bachelors_degree, above_bachelor, total_population)
# Create percentage columns
graph2_df_percentage <- 
  mutate(graph2_df, no_high_school_pt =
           as.numeric(graph2_df$no_high_school)/as.numeric(graph2_df$total_population))
graph2_df_percentage <- 
  mutate(graph2_df_percentage, above_bachelor_pt =
           as.numeric(graph2_df$above_bachelor)/as.numeric(graph2_df$total_population))
graph2_df_percentage <- 
  mutate(graph2_df_percentage, university_diploma_pt =
           as.numeric(graph2_df$university_diploma)/as.numeric(graph2_df$total_population))
graph2_df_percentage <- 
  mutate(graph2_df_percentage, less_than_diploma_pt =
           as.numeric(graph2_df$less_than_diploma)/as.numeric(graph2_df$total_population))
graph2_df_percentage <- 
  mutate(graph2_df_percentage, bachelors_degree_pt =
           as.numeric(graph2_df$bachelors_degree)/as.numeric(graph2_df$total_population))
graph2_df_percentage <- graph2_df_percentage %>%
  select(
    no_high_school_pt,
    above_bachelor_pt,
    university_diploma_pt,
    less_than_diploma_pt,
    bachelors_degree_pt,
    neighbourhood
  )
#graph2_df_percentage
#Reshape the dataset
graph2_df_percentage_long <- 
  graph2_df_percentage %>% 
  pivot_longer(cols = no_high_school_pt:bachelors_degree_pt, names_to = "study_level")
#graph2_df_percentage_long
```

## Plotting Graph2

```{r echo= FALSE}
ggplot(
graph2_df_percentage_long, aes(x = neighbourhood, y = value, fill = study_level)) + 
  # Specify that we want a bar graph:
  geom_bar(stat="identity", position=position_dodge()) +
  #geom_text(aes(y=label_ypos, label=c('Bridle Path - Sunnybrook - York Mills', 'Regent Park', 'Toronto'))) +
  # Add titles and axis labels:
  labs(title = "Neighborhoods by study level", 
       x = "Neighborhood", 
       y = "Population") 
```
