---
title: "Demography of Education in Toronto's neighbourhoods with Highest and Lowest Income"
date: "23/01/2020"
author: Ke-Li Chiu, Diego Mamanche Castellanos
output: pdf_document
fontsize: 5pt

abstract: High levels of education, the common characteristics of those who live in Bridle Path-Sunnybrook-York Mills, Toronto's neighbourhood found as the one with the highest median income. By using Toronto's open data, this paper illustrates the education level distribution of the highest and lowest neighbourhood in terms of household income. It is important to understand these distributions because it exposes an alarming income inequality, that can be associated with the education level in each family.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r, echo=FALSE, include=FALSE}
# installing packages
my_packages <- c("opendatatoronto", "dplyr", "ggplot2", "tidyr", "captioner")
# For each of the 3 packages, check whether each one exists, and install if not
for (pkg in my_packages) {
  
  if (!pkg %in% installed.packages()) { 
    # If package is not installed, install it:
    install.packages(pkg)
  } 
}
# Importing libraries
library(opendatatoronto)
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
```

```{r, echo= FALSE}
# Search packages (this returns a table)
wellbeing_packages <- search_packages("Wellbeing")
# Print the table
#wellbeing_packages
```

```{r, echo= FALSE}
### Filtering the wellbeing demographic dataset
wellbeing_demographics_package<- wellbeing_packages %>%
  filter(title == "Wellbeing Toronto - Demographics: NHS Indicators") # Only keep the row(s) where the title is "Wellbeing Toronto - Demographics: NHS Indicators"
```

```{r, echo= FALSE}
# Getting the information dataset called "demographics-nhs-indicators-2011"
demographic_indicators2011 <- wellbeing_demographics_package %>% # Start with the package 
  list_package_resources() %>% # List the resources in the package
  filter(name == "demographics-nhs-indicators-2011") %>% # Only keep the resource we want 
  get_resource()
# kable(head(demographic_indicators2011[,c(1,47:48)])) # Show a preview of the data set
```

```{r, echo=FALSE}
#Cleaning the column names using janitor::clean_names
demographic_indicators2011 <-
  janitor::clean_names(demographic_indicators2011)
```

```{r, echo= FALSE}
#Selecting columns needed
income_demographic2011 <- subset(demographic_indicators2011, select = c("hood_number", "median_after_tax_household_income","median_household_total_income"))
names(income_demographic2011)[names(income_demographic2011) == "hood_number"] <- "ID"
#income_demographic2011  <- as.data.frame(income_demographic2011)
income_demographic2011 <- income_demographic2011[income_demographic2011$ID != 'ID',]       
#income_demographic2011
```


```{r, echo= FALSE, message=FALSE}
education_indicators2011 <- wellbeing_demographics_package %>% # Start with the package 
  list_package_resources() %>% # List the resources in the package
  filter(name == "education-nhs-indicators-2011") %>% # Only keep the resource we want 
  get_resource()
```


```{r, echo= FALSE, message=FALSE}
education_indicators2011 <- wellbeing_demographics_package %>% # Start with the package 
  list_package_resources() %>% # List the resources in the package
  filter(name == "education-nhs-indicators-2011") %>% # Only keep the resource we want 
  get_resource()
```

```{r, echo= FALSE}
#Cleaning the column names using janitor::clean_names
education_indicators2011 <-
  janitor::clean_names(education_indicators2011)
#head(education_indicators2011[,1:4])
```


```{r echo= FALSE}
#Extracting the neighbourhood code from the column "geography"
edu_id <- extract(education_indicators2011, geography, into = "ID" , regex = "([(][0-9]+[)])")
edu_id <- extract(edu_id, ID, into = "ID" , regex = "([0-9]+)")
#head(edu_id[,1:4])
```

```{r, echo= FALSE}
###Filterint the housing dataset in order to extract the names of each neighbourhood.
wellbeing_housing_package <- wellbeing_packages %>%
  filter(title == "Wellbeing Toronto - Housing") 
#Getting the information dataset called "wellbeing-toronto-housing"
housing_indicators2011 <- wellbeing_housing_package %>% # Start with the package 
  list_package_resources() %>% # List the resources in the package
  filter(name == "wellbeing-toronto-housing") %>% # Only keep the resource we want 
  get_resource()
housing_indicators2011 <- housing_indicators2011$RawDataRef_2011
# kable(housing_indicators2011[1:3,1:2])
```

```{r echo= FALSE}
#Cleaning the column names using janitor::clean_names, and selecting columns needed
housing_indicators2011 <- 
  janitor::clean_names(housing_indicators2011)
housing_indicators2011 <- subset(housing_indicators2011, select = c("neighbourhood_id", "neighbourhood"))
names(housing_indicators2011)[names(housing_indicators2011) == "neighbourhood_id"] <- "ID"
#housing_indicators2011
```

```{r echo= FALSE, include=FALSE, message=FALSE}
merged_df <- merge(housing_indicators2011, income_demographic2011, by = 'ID')
merged_df <- merge(merged_df, edu_id, by = 'ID')
merged_df <- merged_df[,1:16]
#head(merged_df)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
#CLEANING THE MERGED DATASET
#Calculate the median for column "median_after_tax_household_income"
#Note: In this case, the median is better due to outliers
  
toronto_median_income <- median(as.numeric(income_demographic2011$median_after_tax_household_income))
#Calculate mean for the rest of the columns
toronto_other_columns <-  mutate_all(merged_df[,4:16], as.numeric)
toronto_other_columns <- colMeans(toronto_other_columns)
toronto_other_columns <- as.data.frame(toronto_other_columns)
toronto_other_columns <- round(toronto_other_columns[,1])
#Identify max and min neighbourhoods
neighbourhoods <- c(merged_df[which.max(merged_df$median_after_tax_household_income),]$neighbourhood, merged_df[which.min(merged_df$median_after_tax_household_income),]$neighbourhood)
#Create graph2 dataframe
filtered_dataset <- filter(merged_df, neighbourhood %in% c(neighbourhoods))
filtered_dataset <- add_row(filtered_dataset)
#Assign values to third row
filtered_dataset[3,2:3] <- c('Toronto', toronto_median_income)
filtered_dataset[3,4:16] <- toronto_other_columns
#Create column less_than_diploma
filtered_dataset <- mutate(filtered_dataset, less_than_diploma = high_school_diploma_or_equivalent + 
apprenticeship_or_trades_certificate_or_diploma + 
college_cegep_or_other_non_university_certificate_or_diploma)
#Rename columns
filtered_dataset <- 
  rename(filtered_dataset, 
    median_income = median_household_total_income,
    median_income_after_tax = median_after_tax_household_income,
    no_high_school = no_certificate_diploma_or_degree,
    above_bachelor = university_certificate_diploma_or_degree_at_bachelor_level_or_above,
    university_diploma = university_certificate_or_diploma_below_bachelor_level
    )
#Create column total_population
filtered_dataset <- mutate(filtered_dataset, total_population = (    no_high_school + above_bachelor + university_diploma + less_than_diploma + bachelors_degree ))
#Print the table
#kable(head(filtered_dataset[,1:4]))
```

```{r echo= FALSE}
#Create the dataframe graph1_df to plot the values
graph1_df <- select(filtered_dataset, neighbourhood, median_income_after_tax, median_income)
#Reshape the dataset
graph1_df_reshaped <- gather(graph1_df, "income_type", "income_value", -1)
#graph1_df
```



```{r echo= FALSE}
#Create the dataframe graph2_df to plot the values
graph2_df <- select(filtered_dataset, neighbourhood, no_high_school, less_than_diploma, university_diploma, bachelors_degree, above_bachelor, total_population)
# Create percentage columns
graph2_df_percentage <- 
  mutate(graph2_df, no_high_school_pt =
           round(as.numeric(graph2_df$no_high_school)/as.numeric(graph2_df$total_population)*100), digits=4)
graph2_df_percentage <- 
  mutate(graph2_df_percentage, above_bachelor_pt =
            round(as.numeric(graph2_df$above_bachelor)/as.numeric(graph2_df$total_population)*100), digits=4)
graph2_df_percentage <- 
  mutate(graph2_df_percentage, university_diploma_pt =
            round(as.numeric(graph2_df$university_diploma)/as.numeric(graph2_df$total_population)*100), digits=4)
graph2_df_percentage <- 
  mutate(graph2_df_percentage, less_than_diploma_pt =
            round(as.numeric(graph2_df$less_than_diploma)/as.numeric(graph2_df$total_population)*100), digits=4)
graph2_df_percentage <- 
  mutate(graph2_df_percentage, bachelors_degree_pt =
            round(as.numeric(graph2_df$bachelors_degree)/as.numeric(graph2_df$total_population)*100), digits=4)
graph2_df_percentage <- graph2_df_percentage %>%
  select(
    neighbourhood,
    no_high_school_pt,
    above_bachelor_pt,
    university_diploma_pt,
    less_than_diploma_pt,
    bachelors_degree_pt,
  )
#Reshape the dataset
graph2_df_percentage_long <- 
  graph2_df_percentage %>% 
  pivot_longer(cols = no_high_school_pt:bachelors_degree_pt, names_to = "study_level")
#graph2_df_percentage_long
```

# Wellbeing Toronto - Demographics: NHS Indicators
The data set package we delve into is *Wellbeing Toronto - Demographics: NHS Indicators*, which provides information from the 2011 National Household Survey (NHS) and presents a profile of Toronto. Data is broken down by the 140 neighbourhoods that make up the City of Toronto. With this data set package, we first identify the two neighbourhoods that have the highest income and the lowest income and then analysize the education demographics in these two neighbourhoods to understand difference of education attainment between these two neighbourhood. We also identify the median household income and the average education levels of the city to be compared with the two neighbourhoods. There are two data sets in this package used in the analysis: _Demographics NHS Indicators 2011_ and _Education NHS Indicators 2011_. Additionally, we incorporate the *Wellbeing Toronto - Housing* data set solely for extracting the names of the neighbourhoods; no housing indicators are used in our analysis. 

## Demographics NHS Indicators 2011 data set
_Demographics NHS Indicators 2011_ shows the profile of each neighbourhood regarding population breakdown by ethnicity, household income, education level, employment, age and sex. The indicators we are interested in are *Median household total income* and *Median after-tax household income*. These two columns show the median household income of every neighbourhood before and after tax in Canadian dollars. 

From this data set, we identify the neighbourd with the highest median household income (Bridle Path-Sunnybrook-York Mills) and the neighbourd with the lowest median household income (Regent Park). We also calculate the median household income of all neibourhoods in the city of Toronto as summary statistics to be compared with the two neighbourhoods, appended to the last row.

Table  1: Median household income before and after tax.
```{r echo= FALSE}
kable(graph1_df)
```

Graph 1: Median Total Household Income Graph

```{r echo= FALSE, fig1, fig.height = 2, fig.width = 7}
#Plotting the graph
ggplot(graph1_df_reshaped, aes(x = neighbourhood, y = as.numeric(income_value), fill = income_type)) + 
  # Specify that we want a bar graph:
  geom_bar(stat="identity", position=position_dodge()) + 
  # Add titles and axis labels:
  labs(title = 
         "Neighbourhoods with the Highest and Lowest median income, \nand Toronto median income", x = "neighbourhood", y = "Income") 
```


## Education NHS Indicators 2011
_Education NHS Indicators_ is provided for the population of age 15 and over living in private households. The data set presents the distrubution of people's most advanced education level in each neighbourhood. Because the population in each neighbourhood is different, we can not compare the percentage of education attainment by the absolute values provided in the data set. Therefore, we transform the data to relative values. The transformation is done by aggregating the number of population in all education levels and dividing it by the number in each education level. 
```{r echo= FALSE}
kable(graph2_df_percentage[,1:4])
```


Graph 2: Education Levels Percentage

```{r echo= TRUE, fig2, fig.height = 3, fig.width = 9}
#Plotting the graph
ggplot(graph2_df_percentage_long, aes(x = neighbourhood, y = value, fill = study_level)) + 
  # Specify that we want a bar graph:
  geom_bar(stat="identity", position=position_dodge()) +
  # Add titles and axis labels:
  labs(title = "Neighborhoods by study level", x = "Neighborhood", 
       y = "Population (%)") 
```
