---
title: "Compairing Toronto neighbourhoods With The Highest And Lowest Income"
authors: "Ke-li Chiu & Diego Mamanche Castellanos"
date: "23/01/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=FALSE}
# installing packages
my_packages <- c("opendatatoronto", "dplyr", "ggplot2", "tidyr")
# For each of the 3 packages, check whether each one exists, and install if not
for (pkg in my_packages) {
  
  if (!pkg %in% installed.packages()) { 
    # If package is not installed, install it:
    install.packages(pkg)
  } 
}
# Importing libraries
library(opendatatoronto)
library(dplyr)
library(tidyr)
library(ggplot2)
```

# ABOUT THE DATA


```{r, echo= FALSE}
# Search packages (this returns a table)
wellbeing_packages <- search_packages("Wellbeing")
# Print the table
#wellbeing_packages
```



## The Wellbeing Toronto datasets

Explanation



```{r, echo= FALSE}
### Filtering the wellbeing demographic dataset
wellbeing_demographics_package<- wellbeing_packages %>%
  filter(title == "Wellbeing Toronto - Demographics: NHS Indicators") # Only keep the row(s) where the title is "Wellbeing Toronto - Demographics: NHS Indicators"
```


List the names of the internal datasets

```{r, echo= FALSE}
list_package_resources(wellbeing_demographics_package)
```

## First dataset

We chose demographics-nhs-indicators-2011 internal dataset for information about income

```{r, echo= FALSE}
#Getting the information dataset called "demographics-nhs-indicators-2011"
demographic_indicators2011 <- wellbeing_demographics_package %>% # Start with the package 
  list_package_resources() %>% # List the resources in the package
  filter(name == "demographics-nhs-indicators-2011") %>% # Only keep the resource we want 
  get_resource()
```



```{r, echo=FALSE}
#Cleaning the column names using janitor::clean_names
demographic_indicators2011 <-
  janitor::clean_names(demographic_indicators2011)
head(demographic_indicators2011[,c(1,46:48)])
```



```{r, echo= FALSE}
#Selecting columns needed
income_demographic2011 <- subset(demographic_indicators2011, select = c("hood_number","total_population", "median_after_tax_household_income","median_household_total_income", 'total_visible_minority_population', 'not_a_visible_minority'))
names(income_demographic2011)[names(income_demographic2011) == "hood_number"] <- "ID"
#income_demographic2011  <- as.data.frame(income_demographic2011)
income_demographic2011 <- income_demographic2011[income_demographic2011$ID != 'ID',]       
#income_demographic2011
```

##Second dataset

We chose education-nhs-indicators-2011 internal dataset for information about education levels 


```{r, echo= FALSE}
education_indicators2011 <- wellbeing_demographics_package %>% # Start with the package 
  list_package_resources() %>% # List the resources in the package
  filter(name == "education-nhs-indicators-2011") %>% # Only keep the resource we want 
  get_resource()
#education_indicators2011
```




```{r, echo=TRUE}
#Cleaning the column names using janitor::clean_names
education_indicators2011 <-
  janitor::clean_names(education_indicators2011)
head(education_indicators2011[,1:4])
```

We extract the code from column "geography" and the levels of education.

```{r}
#Extracting the neighbourhood code from the column "geography"
edu_id <- extract(education_indicators2011, geography, into = "ID" , regex = "([(][0-9]+[)])")
edu_id <- extract(edu_id, ID, into = "ID" , regex = "([0-9]+)")
head(edu_id[,1:4])
```


## Third dataset

We chose "Wellbeing Toronto - Housing" dataset in order to extract the names of the neighbourhood.

```{r, echo= TRUE}
###Filterint the housing dataset in order to extract the names of each neighbourhood.
wellbeing_housing_package <- wellbeing_packages %>%
  filter(title == "Wellbeing Toronto - Housing") # Only keep the row(s) where the title is "Wellbeing Toronto - Demographics: NHS Indicators"
```




```{r, echo= TRUE}
#Extracting the package
list_package_resources(wellbeing_housing_package)
```



```{r, echo= TRUE}
#Getting the information dataset called "wellbeing-toronto-housing"
housing_indicators2011 <- wellbeing_housing_package %>% # Start with the package 
  list_package_resources() %>% # List the resources in the package
  filter(name == "wellbeing-toronto-housing") %>% # Only keep the resource we want 
  get_resource()
housing_indicators2011 <- housing_indicators2011$RawDataRef_2011
head(housing_indicators2011[,1:4])
```




```{r}
#Cleaning the column names using janitor::clean_names, and selecting columns needed
housing_indicators2011 <- 
  janitor::clean_names(housing_indicators2011)
housing_indicators2011 <- subset(housing_indicators2011, select = c("neighbourhood_id", "neighbourhood"))
names(housing_indicators2011)[names(housing_indicators2011) == "neighbourhood_id"] <- "ID"
#housing_indicators2011
```




## Merging all datasets

Explanation

```{r}
merged_df <- merge(housing_indicators2011, income_demographic2011, by = 'ID')
merged_df <- merge(merged_df, edu_id, by = 'ID')
merged_df <- merged_df[,1:16]
head(merged_df)
```

```{r echo=FALSE}
#CLEANING THE MERGED DATASET
#Calculate the median for column "median_after_tax_household_income"
#Note: In this case, the median is better due to outliers
  
toronto_median_income <- median(as.numeric(income_demographic2011$median_after_tax_household_income))
toronto_median_income
#Calculate mean for the rest of the columns
toronto_other_columns <-  mutate_all(merged_df[,4:16], as.numeric)
toronto_other_columns <- colMeans(toronto_other_columns)
toronto_other_columns <- as.data.frame(toronto_other_columns)
toronto_other_columns <- round(toronto_other_columns[,1])
toronto_other_columns
neighbourhoods <- c(merged_df[which.max(merged_df$median_after_tax_household_income),]$neighbourhood, merged_df[which.min(merged_df$median_after_tax_household_income),]$neighbourhood)
#Create graph2 dataframe
filtered_dataset <- filter(merged_df, neighbourhood %in% c(neighbourhoods))
filtered_dataset <- add_row(filtered_dataset)
#Assign values to third row
filtered_dataset[3,2:3] <- c('Toronto', toronto_median_income)
filtered_dataset[3,4:16] <- toronto_other_columns
filtered_dataset
filtered_dataset <- mutate(filtered_dataset, less_than_diploma = high_school_diploma_or_equivalent + 
postsecondary_certificate_diploma_or_degree + 
apprenticeship_or_trades_certificate_or_diploma + 
college_cegep_or_other_non_university_certificate_or_diploma)
#Rename columns
filtered_dataset <- 
  rename(filtered_dataset, 
    median_income = median_household_total_income,
    median_income_after_tax = median_after_tax_household_income,
    no_high_school = no_certificate_diploma_or_degree,
    above_bachelor = university_certificate_diploma_or_degree_at_bachelor_level_or_above,
    university_diploma = university_certificate_or_diploma_below_bachelor_level
    )
head(filtered_dataset)
```


#Create Graph1 dataframe 

```{r}
#Create the dataframe graph1_df to plot the values
graph1_df <- select(filtered_dataset, neighbourhood, median_income_after_tax, median_income)
#Reshape the dataset
graph1_df <- gather(graph1_df, "income_type", "income_value", -1)
graph1_df
```


Plotting the graph

```{r}
ggplot(graph1_df, aes(x = neighbourhood, y = income_value, fill = income_type)) + 
  # Specify that we want a bar graph:
  geom_bar(stat="identity", position=position_dodge()) + 
  # Add titles and axis labels:
  labs(title = "neighbourhoods with the Highest and Lowest median income, \nand Toronto median income", 
       x = "neighbourhood", 
       y = "Income") 
```

##Create graph2 dataframe

```{r}
#Create the dataframe graph1_df to plot the values
graph2_df <- select(filtered_dataset, neighbourhood, no_high_school, less_than_diploma, university_diploma, bachelors_degree, above_bachelor, total_population)

# Create percentage columns
graph2_df_percentage <- 
  mutate(graph2_df, no_high_school_pt =
           as.numeric(graph2_df$no_high_school)/as.numeric(graph2_df$total_population))
graph2_df_percentage <- 
  mutate(graph2_df_percentage, above_bachelor_pt =
           as.numeric(graph2_df$above_bachelor)/as.numeric(graph2_df$total_population))
graph2_df_percentage <- 
  mutate(graph2_df_percentage, university_diploma_pt =
           as.numeric(graph2_df$university_diploma)/as.numeric(graph2_df$total_population))
graph2_df_percentage <- 
  mutate(graph2_df_percentage, less_than_diploma_pt =
           as.numeric(graph2_df$less_than_diploma)/as.numeric(graph2_df$total_population))
graph2_df_percentage <- 
  mutate(graph2_df_percentage, bachelors_degree_pt =
           as.numeric(graph2_df$bachelors_degree)/as.numeric(graph2_df$total_population))
graph2_df_percentage <- graph2_df_percentage %>%
  select(
    no_high_school_pt,
    above_bachelor_pt,
    university_diploma_pt,
    less_than_diploma_pt,
    bachelors_degree_pt,
    neighbourhood
  )
graph2_df_percentage


#Reshape the dataset
graph2_df_percentage_long <- 
  graph2_df_percentage %>% 
  pivot_longer(cols = no_high_school_pt:bachelors_degree_pt, names_to = "study_level")

graph2_df_percentage_long
```

##Plotting Graph2

```{r}
ggplot(
graph2_df_percentage_long, aes(x = neighbourhood, y = value, fill = study_level)) + 
  # Specify that we want a bar graph:
  geom_bar(stat="identity", position=position_dodge()) +
  #geom_text(aes(y=label_ypos, label=c('Bridle Path - Sunnybrook - York Mills', 'Regent Park', 'Toronto'))) +
  # Add titles and axis labels:
  labs(title = "Neighborhoods by study level", 
       x = "Neighborhood", 
       y = "Population") 
```
