---
title: "Analysing Education Level in Toronto by Neighborhood"
author: "Ke-li Chiu & Diego Mamanche Castellanos"
date: "23/01/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Ke-li's writing!!!!!!

Ke-li 


Importing libraries

```{r, echo=TRUE}
my_packages <- c("opendatatoronto", "dplyr", "ggplot2", "tidyr")

# For each of the 3 packages, check whether each one exists, and install if not
for (pkg in my_packages) {
  
  if (!pkg %in% installed.packages()) { 
    # If package is not installed, install it:
    install.packages(pkg)
  } 
}

# Importing libraries
library(opendatatoronto)
library(dplyr)
library(tidyr)
library(ggplot2)

```

## Guess what, more Keli's writing!!!

## Finding the data

To find a dataset you like, you have a couple options:

1. Search for a dataset manually via the [Toronto Open Data Portal](https://open.toronto.ca/)
2. Use the `search_packages()` function

Here we'll use the `search_packages()` function to keep this document self-contained.


```{r, echo= TRUE}
# Search packages (this returns a table)
wellbeing_packages <- search_packages("Wellbeing")

# Print the table
delay_packages

```



You guessed right!!!! more Ke-li´s writing.

```{r, echo= TRUE}
wellbeing_demographics_package<- wellbeing_packages %>%
  filter(title == "Wellbeing Toronto - Demographics: NHS Indicators") # Only keep the row(s) where the title is "Wellbeing Toronto - Demographics: NHS Indicators"

```


More writing

```{r, echo= TRUE}
list_package_resources(wellbeing_demographics_package)

```

Writing

```{r, echo= TRUE}
demographic_indicators2011 <- wellbeing_demographics_package %>% # Start with the package 
  list_package_resources() %>% # List the resources in the package
  filter(name == "demographics-nhs-indicators-2011") %>% # Only keep the resource we want 
  get_resource()
```


We have the data

```{r, echo=TRUE}
demographic_indicators2011
```


Writing


```{r}
# See what the "Line" column looks like
edu_demographic2011 = demographic_indicators2011 %>% select('College, CEGEP or other non-university certificate or diploma')

edu_demographic2011 <- edu_demographic2011 %>% 
  rename(
    'Less than Bachelor' = 'College, CEGEP or other non-university certificate or diploma'
    )

edu_demographic2011 <- edu_demographic2011 %>%
  dplyr::slice(2:)

edu_demographic2011

```
