---
title: "Compairing Toronto Neighborhoods With The Highest And Lowest Income"
author: "Ke-li Chiu & Diego Mamanche Castellanos"
date: "23/01/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Ke-li's writing!!!!!!

Ke-li 


Importing libraries

```{r, echo=TRUE}
my_packages <- c("opendatatoronto", "dplyr", "ggplot2", "tidyr")

# For each of the 3 packages, check whether each one exists, and install if not
for (pkg in my_packages) {
  
  if (!pkg %in% installed.packages()) { 
    # If package is not installed, install it:
    install.packages(pkg)
  } 
}

# Importing libraries
library(opendatatoronto)
library(dplyr)
library(tidyr)
library(ggplot2)

```

## Guess what, more Keli's writing!!!

## Finding the data

To find a dataset you like, you have a couple options:

1. Search for a dataset manually via the [Toronto Open Data Portal](https://open.toronto.ca/)
2. Use the `search_packages()` function

Here we'll use the `search_packages()` function to keep this document self-contained.


```{r, echo= TRUE}
# Search packages (this returns a table)
wellbeing_packages <- search_packages("Wellbeing")

# Print the table
wellbeing_packages

```



You guessed right!!!! more Ke-li´s writing.

```{r, echo= TRUE}
wellbeing_demographics_package<- wellbeing_packages %>%
  filter(title == "Wellbeing Toronto - Demographics: NHS Indicators") # Only keep the row(s) where the title is "Wellbeing Toronto - Demographics: NHS Indicators"

```


More writing

```{r, echo= TRUE}
list_package_resources(wellbeing_demographics_package)

```

Writing

```{r, echo= TRUE}
demographic_indicators2011 <- wellbeing_demographics_package %>% # Start with the package 
  list_package_resources() %>% # List the resources in the package
  filter(name == "demographics-nhs-indicators-2011") %>% # Only keep the resource we want 
  get_resource()
```





We have the data

```{r, echo=TRUE}
demographic_indicators2011 <-
  janitor::clean_names(demographic_indicators2011)

demographic_indicators2011
```

```{r, echo= TRUE}
income_demographic2011 <- subset(demographic_indicators2011, select = c("hood_number", "median_after_tax_household_income", 'no_certificate_diploma_or_degree', 'college_cegep_or_other_non_university_certificate_or_diploma', 'university_certificate_or_diploma_below_bachelor_level', 'total_visible_minority_population', 'not_a_visible_minority'))

names(income_demographic2011)[names(income_demographic2011) == "hood_number"] <- "ID"

#income_demographic2011  <- as.data.frame(income_demographic2011)

income_demographic2011 <- income_demographic2011[income_demographic2011$ID != 'ID',]       

income_demographic2011
```




Second dataset


```{r, echo= TRUE}
wellbeing_housing_package <- wellbeing_packages %>%
  filter(title == "Wellbeing Toronto - Housing") # Only keep the row(s) where the title is "Wellbeing Toronto - Demographics: NHS Indicators"

```


More writing

```{r, echo= TRUE}
list_package_resources(wellbeing_housing_package)

```

Writing

```{r, echo= TRUE}
housing_indicators2011 <- wellbeing_housing_package %>% # Start with the package 
  list_package_resources() %>% # List the resources in the package
  filter(name == "wellbeing-toronto-housing") %>% # Only keep the resource we want 
  get_resource()

housing_indicators2011 <- housing_indicators2011$RawDataRef_2011
```


We have the data

```{r, echo=TRUE}
housing_indicators2011
```

```{r}
housing_indicators2011 <- 
  janitor::clean_names(housing_indicators2011)

housing_indicators2011 <- subset(housing_indicators2011, select = c("neighbourhood_id", "neighbourhood"))

names(housing_indicators2011)[names(housing_indicators2011) == "neighbourhood_id"] <- "ID"


housing_indicators2011
```

```{r}
merged_df <- merge(housing_indicators2011, income_demographic2011, by = 'ID')

merged_df
```


```{r}
#Create Graph1 dataframe 

#income_demographic2011 <- transform(income_demographic2011, Median_income #= as.numeric(Median_income))


values <- c(median(as.numeric(merged_df$median_after_tax_household_income)), max(as.numeric(merged_df$median_after_tax_household_income)), min(as.numeric(merged_df$median_after_tax_household_income)))

neighborhoods <- c('Toronto Median', merged_df[which.max(merged_df$median_after_tax_household_income),]$neighbourhood, merged_df[which.min(merged_df$median_after_tax_household_income),]$neighbourhood)

measures <- c('median', 'max', 'min')

graph1_df <- cbind.data.frame(measures, neighborhoods, values)
graph1_df

```


```