---
title: "Income Gap and Education Profile in Toronto's Neighbourhoods"
date: "23/01/2020"
author: Ke-Li Chiu, Diego Mamanche Castellanos
output: pdf_document
geometry: margin=1in
fontsize: 5pt

abstract: "Education is a contributor to many beneficial socio-economic outcomes. We intend to exam the education profile in the selected neighbourhoods of Toronto that have polarized household income from the data set package *Wellbeing Toronto - Demographics: NHS Indicators*. We first observe that the income gap between the neighbourhood with the highest household income and the lowest is enormous. The median household income of Bridle Path-Sunnybrook-York Mills is almost nine times more than Regent Park before tax. Bridle Path-Sunnybrook-York Mills also has a much higher attainment rate (45%) of higher education above a bachelor's degree than Regent Park (23%) and the City of Toronto (27%). We intend to use the comparison to bring awareness of inequality in income and education existing in the City of Toronto."
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r, echo=FALSE, include=FALSE}
# installing packages
my_packages <- c("opendatatoronto", "dplyr", "ggplot2", "tidyr", "captioner")
# For each of the 3 packages, check whether each one exists, and install if not
for (pkg in my_packages) {
  
  if (!pkg %in% installed.packages()) { 
    # If package is not installed, install it:
    install.packages(pkg)
  } 
}
# Importing libraries
library(opendatatoronto)
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
```

```{r, echo= FALSE}
# Search packages (this returns a table)
wellbeing_packages <- search_packages("Wellbeing")
# Print the table
#wellbeing_packages
```

```{r, echo= FALSE}
### Filtering the wellbeing demographic dataset
wellbeing_demographics_package<- wellbeing_packages %>%
  filter(title == "Wellbeing Toronto - Demographics: NHS Indicators") # Only keep the row(s) where the title is "Wellbeing Toronto - Demographics: NHS Indicators"
```

```{r, echo= FALSE}
# Getting the information dataset called "demographics-nhs-indicators-2011"
demographic_indicators2011 <- wellbeing_demographics_package %>% # Start with the package 
  list_package_resources() %>% # List the resources in the package
  filter(name == "demographics-nhs-indicators-2011") %>% # Only keep the resource we want 
  get_resource()
# kable(head(demographic_indicators2011[,c(1,47:48)])) # Show a preview of the data set
```

```{r, echo=FALSE}
#Cleaning the column names using janitor::clean_names
demographic_indicators2011 <-
  janitor::clean_names(demographic_indicators2011)
```

```{r, echo= FALSE}
#Selecting columns needed
income_demographic2011 <- subset(demographic_indicators2011, select = c("hood_number", "median_after_tax_household_income","median_household_total_income"))
names(income_demographic2011)[names(income_demographic2011) == "hood_number"] <- "ID"
#income_demographic2011  <- as.data.frame(income_demographic2011)
income_demographic2011 <- income_demographic2011[income_demographic2011$ID != 'ID',]       
#income_demographic2011
```


```{r, echo= FALSE, message=FALSE}
education_indicators2011 <- wellbeing_demographics_package %>% # Start with the package 
  list_package_resources() %>% # List the resources in the package
  filter(name == "education-nhs-indicators-2011") %>% # Only keep the resource we want 
  get_resource()
```


```{r, echo= FALSE, message=FALSE}
education_indicators2011 <- wellbeing_demographics_package %>% # Start with the package 
  list_package_resources() %>% # List the resources in the package
  filter(name == "education-nhs-indicators-2011") %>% # Only keep the resource we want 
  get_resource()
```

```{r, echo= FALSE}
#Cleaning the column names using janitor::clean_names
education_indicators2011 <-
  janitor::clean_names(education_indicators2011)
#head(education_indicators2011[,1:4])
```


```{r echo= FALSE}
#Extracting the neighbourhood code from the column "geography"
edu_id <- extract(education_indicators2011, geography, into = "ID" , regex = "([(][0-9]+[)])")
edu_id <- extract(edu_id, ID, into = "ID" , regex = "([0-9]+)")
#head(edu_id[,1:4])
```

```{r, echo= FALSE}
###Filterint the housing dataset in order to extract the names of each neighbourhood.
wellbeing_housing_package <- wellbeing_packages %>%
  filter(title == "Wellbeing Toronto - Housing") 
#Getting the information dataset called "wellbeing-toronto-housing"
housing_indicators2011 <- wellbeing_housing_package %>% # Start with the package 
  list_package_resources() %>% # List the resources in the package
  filter(name == "wellbeing-toronto-housing") %>% # Only keep the resource we want 
  get_resource()
housing_indicators2011 <- housing_indicators2011$RawDataRef_2011
# kable(housing_indicators2011[1:3,1:2])
```

```{r echo= FALSE}
#Cleaning the column names using janitor::clean_names, and selecting columns needed
housing_indicators2011 <- 
  janitor::clean_names(housing_indicators2011)
housing_indicators2011 <- subset(housing_indicators2011, select = c("neighbourhood_id", "neighbourhood"))
names(housing_indicators2011)[names(housing_indicators2011) == "neighbourhood_id"] <- "ID"
#housing_indicators2011
```

```{r echo= FALSE, include=FALSE, message=FALSE}
merged_df <- merge(housing_indicators2011, income_demographic2011, by = 'ID')
merged_df <- merge(merged_df, edu_id, by = 'ID')
merged_df <- merged_df[,1:16]
#head(merged_df)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
#CLEANING THE MERGED DATASET
#Calculate the median for column "median_after_tax_household_income"
#Note: In this case, the median is better due to outliers
  
toronto_median_income <- median(as.numeric(income_demographic2011$median_after_tax_household_income))
#Calculate mean for the rest of the columns
toronto_other_columns <-  mutate_all(merged_df[,4:16], as.numeric)
toronto_other_columns <- colMeans(toronto_other_columns)
toronto_other_columns <- as.data.frame(toronto_other_columns)
toronto_other_columns <- round(toronto_other_columns[,1])
#Identify max and min neighbourhoods
neighbourhoods <- c(merged_df[which.max(merged_df$median_after_tax_household_income),]$neighbourhood, merged_df[which.min(merged_df$median_after_tax_household_income),]$neighbourhood)
#Create graph2 dataframe
filtered_dataset <- filter(merged_df, neighbourhood %in% c(neighbourhoods))
filtered_dataset <- add_row(filtered_dataset)
#Assign values to third row
filtered_dataset[3,2:3] <- c('Toronto', toronto_median_income)
filtered_dataset[3,4:16] <- toronto_other_columns
#Create column less_than_diploma
filtered_dataset <- mutate(filtered_dataset, less_than_diploma = high_school_diploma_or_equivalent + 
apprenticeship_or_trades_certificate_or_diploma + 
college_cegep_or_other_non_university_certificate_or_diploma)
#Rename columns
filtered_dataset <- 
  rename(filtered_dataset, 
    median_income = median_household_total_income,
    median_income_after_tax = median_after_tax_household_income,
    no_high_school = no_certificate_diploma_or_degree,
    above_bachelor = university_certificate_diploma_or_degree_at_bachelor_level_or_above,
    university_diploma = university_certificate_or_diploma_below_bachelor_level
    )
#Create column total_population
filtered_dataset <- mutate(filtered_dataset, total_population = (    no_high_school + above_bachelor + university_diploma + less_than_diploma + bachelors_degree ))
#Print the table
#kable(head(filtered_dataset[,1:4]))
```

```{r echo= FALSE}
#Create the dataframe graph1_df to plot the values
graph1_df <- select(filtered_dataset, neighbourhood, median_income_after_tax, median_income)
#Reshape the dataset
graph1_df_reshaped <- gather(graph1_df, "income_type", "income_value", -1)
#graph1_df
```

```{r echo= FALSE}
#Create the dataframe graph2_df to plot the values
graph2_df <- select(filtered_dataset, neighbourhood, no_high_school, less_than_diploma, university_diploma, bachelors_degree, above_bachelor, total_population)
# Create percentage columns
graph2_df_percentage <- 
  mutate(graph2_df, no_certificate_pt =
           round(as.numeric(graph2_df$no_high_school)/as.numeric(graph2_df$total_population)*100), digits=4)
graph2_df_percentage <- 
  mutate(graph2_df_percentage, above_bachelor_pt =
            round(as.numeric(graph2_df$above_bachelor)/as.numeric(graph2_df$total_population)*100), digits=4)
graph2_df_percentage <- 
  mutate(graph2_df_percentage, college_pt =
            round(as.numeric(graph2_df$university_diploma)/as.numeric(graph2_df$total_population)*100), digits=4)
graph2_df_percentage <- 
  mutate(graph2_df_percentage, highschool_pt =
            round(as.numeric(graph2_df$less_than_diploma)/as.numeric(graph2_df$total_population)*100), digits=4)
graph2_df_percentage <- 
  mutate(graph2_df_percentage, bachelors_degree_pt =
            round(as.numeric(graph2_df$bachelors_degree)/as.numeric(graph2_df$total_population)*100), digits=4)
graph2_df_percentage <- graph2_df_percentage %>%
  select(
    neighbourhood,
    no_certificate_pt,
    above_bachelor_pt,
    college_pt,
    highschool_pt,
    bachelors_degree_pt,
  )
#Reshape the dataset
graph2_df_percentage_long <- 
  graph2_df_percentage %>% 
  pivot_longer(cols = no_certificate_pt:bachelors_degree_pt, names_to = "study_level")
#graph2_df_percentage_long
```
# Wellbeing Toronto - Demographics: NHS Indicators
*Wellbeing Toronto - Demographics: NHS Indicators* provides information from the 2011 National Household Survey (NHS) and presents a profile of Toronto across 140 neighbourhoods. There are two data sets in this package used in the analysis: _Demographics NHS Indicators 2011_ and _Education NHS Indicators 2011_, described in the following sections.

## Demographics NHS Indicators 2011 data set
In _Demographics NHS Indicators 2011_, the indicators we are interested in are *Median household total income* and *Median after-tax household income*, which show the median household income of every neighbourhood before and after tax is applied. We identify the neighbourhood with the highest median household income (Bridle Path-Sunnybrook-York Mills) and the neighbourhood with the lowest median household income (Regent Park). We then calculate the median household income of Toronto as summary statistics.
```{r echo= FALSE}
kable(graph1_df)
```

Table 1: Median household income before and after tax.

The median total income is presented in a bar graph comparing the two neighbourhoods and the entire city of Toronto. Before tax is applied, Bridle Path-Sunnybrook-York Mills has a median household income that is almost nine times more than Regent Park and three times more than the median household income of Toronto. We observe that taxation brings the income gap smaller; after tax is applied, the income of Bridle Path-Sunnybrook-York Mills is about five times more than Regent Park.
```{r echo= FALSE, fig1, fig.height = 2, fig.width = 7}
#Plotting the graph
ggplot(graph1_df_reshaped, aes(x = neighbourhood, y = as.numeric(income_value), fill = income_type)) + 
  # Specify that we want a bar graph:
  geom_bar(stat="identity", position=position_dodge()) + 
  # Add titles and axis labels:
  labs(title = "Median Total Household Income", x = "Neighbourhood", y = "Income")+
  theme(text = element_text(size=9)) +
  scale_fill_discrete(name = "", labels = c("Before Tax", "After Tax"))
```
Figure 1: Median Total Household Income Graph

## Education NHS Indicators 2011
The data set presents the distribution of people’s most advanced education level in each neighbourhood. The value provided is measured in the number of people, thus not suitable for percentage comparison. We transform the data to relative values as population percentage by aggregating the number of population in all education levels and dividing it by the population in each education level. We also calculate the average percentage of education levels in Toronto as summary statistics.
```{r echo= FALSE}
kable(graph2_df_percentage[,1:4])
```

Table 2: Education levels percentage

The education level percentages are presented in a bar graph. We observe that Bridle Path-Sunnybrook-York Mills, the neighbourhood with the highest income, has much higher attainment to education above bachelor's degree (45%) compared to Regent Park (23%) and the city of Toronto (27%).
```{r echo= TRUE, fig2, fig.height = 2, fig.width = 9}
ggplot(graph2_df_percentage_long, aes(x = neighbourhood, y = value, fill = study_level)) + 
  # Specify that we want a bar graph:
  geom_bar(stat="identity", position=position_dodge()) +
  # Add graph title and axis labels:
  labs(title = "Education Levels Population Percentage", x = "Neighborhood", 
       y = "Population (%)") +
  # Rename the legend labels to be more readable
  scale_fill_discrete(name = "", labels = c("Above Bachelor", "Bachelor", "Less than Bachelor", 
                                            "Highschool", "No Certificate"))
```
Figure 2: Education Levels Population Percentage Graph


# References

City of Toronto. (2017, November 14). About Wellbeing Toronto. Retrieved January 22, 2020, from https://www.toronto.ca/city-government/data-research-maps/neighbourhoods-communities/wellbeing-toronto/about-wellbeing-toronto/

Open Data Dataset. (n.d.). Retrieved January 22, 2020, from https://open.toronto.ca/dataset/wellbeing-toronto-demographics-nhs-indicators/
